---
output: github_document
bibliography: "inst/REFERENCES.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `SimRescueMeds`

<!-- badges: start -->
<!-- badges: end -->

The goal of `SimRescueMeds` is to enable (relatively) simple simulation studies that impose rescue medication (or really any intercurrent event) on longitudinal data.  

1. Longitudinal data is simulated according a linear mixed model. This is the data one would observe if no intercurrent event, e.g., rescue therapy, occurs.
2. Rescue medication/intercurrent events are simulated using models adapted from @Thomadakis2019, which allows for reasonably straightforward generation of events that, if excluding post-event data, produce MCAR, MAR, or MNAR missingness on the longitudinal data using either exponential or Weibull models. 
3. We also intend to develop functions that can modify post-event data.  

## Installation

You can install the development version of `SimRescueMeds` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jmleach-bst/SimRescueMeds")
```

## Examples

First, load the package.

```{r example}
library(SimRescueMeds)
```

Use `build_design_matrix()` to build the design matrix for the fixed effects, in which you specify the number of individuals in each group `N_t` and `N_c`, the number of time points `K`, the starting time, and time scale. The example below scales such that the file time point has $t = 1$, e.g., 1-year follow-up.

```{r}
xmat_t <- build_design_matrix(
    N_t = 1,
    N_c = 0,
    K = 5,
    time_start = 0,
    time_scale = 1/4
  )
xmat_t
```

Use `simulate_lmm_rct()` to generate data sets based on the linear mixed model, which generates data from a multivariate Normal distribution. Here, we specify a single-individual design matrix for each of 2 treatment groups and the number of individuals in each group. Use `build_lmm_cov()` to more easily construct the covariance matrix as a function of random effects (co)variances and within-individual random error covariances.

```{r}
set.seed(3002242)
df <- simulate_lmm_rct(
  N_t = 50,
  N_c = 50,
  xmat_t = xmat_t,
  xmat_c = build_design_matrix(
    N_t = 0,
    N_c = 1,
    K = 5,
    time_start = 0,
    time_scale = 1/4
  ),
  betas = c(
    beta0 = 8,
    beta1 = 0,
    beta2 = -1,
    beta3 = -1
  ),
  Sigma = build_lmm_cov(
    zmat = xmat_t[, c("intercept", "time")],
    re_sigma = c(1.15, 1.05),
    re_corr_mat = matrix(
      c(1, 0,
        0, 1),
      byrow = TRUE,
      nrow = 2
      ),
    ws_sigma = 1.25,
    ws_corr_mat = diag(5),
    print_intermediate = FALSE
    )
)

head(df, 10)
```

Simulate rescue therapy based on a proportional hazards model in the manner of @Thomadakis2019, which allows the hazard of rescue therapy within some interval $t_{i,j} \le t < t_{i,j+1}$ to be a function of the last observed outcome measurement, $y_{i,j}$, and the next measurement, $y_{i,j+1}$, the latter of which is unobserved in the study since it occurs after rescue therapy. While simple in some respects, this approach allows us to easily define MCAR $(\theta_1 = \theta_2 = 0)$, MAR $(\theta_1 \ne 0,\,\theta_2 = 0)$, or MNAR $(\theta_2 \ne 0)$. 

$$
h_i(t) = \lambda\exp\left(y_{i,j}\theta_1 + y_{i,j+1}\theta_2\right), \quad t_{i,j} \le t < t_{i,j+1}.
$$


```{r}
set.seed(707582)

rt <- sim_hazard_thomadakis_df(
  data = df,
  theta1 = 1/3,
  theta2 = 0,
  lambda = 0.05,
  dist = "exponential",
  print_intermediate = FALSE
)

head(rt)
```

Finally, for analysis we would exclude the events that occur post-event using `exclude_post_tte()`.

```{r}
post_rt <- exclude_post_tte(
  original_data = df,
  tte_data = rt
)

head(post_rt, 15)
```

# References {-}
